RESEARCH IN CONTEXT: Seed ecology of European mesic meadows: the domestication of plant regeneration strategies

### 2021-07-16 README.Rmd
submitted AoB 🌱

# Seed germination ecology of European mesic grasslands

This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

>>>>

# Seed ecology of European mesic meadows: the domestication of plant regeneration strategies

This repository stores all the information related to the manuscript *Seed ecology of European mesic meadows: the domestication of plant regeneration strategies*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

### 2021-07-16 doc/manuscript.Rmd
submitted AoB 🌱

---
title: "Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

@@

Eduardo Fernández-Pascual^1^ ^+^, Madalena Vaz^2^, Beatriz Morais^1^, Ramón Reiné^3^, Joaquín Ascaso^3^, Elías Afif Khouri^1^, Angelino Carta^4^

^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^3^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^4^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy

^+^ Correspondence: Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain. Email: efernandezpascual@gmail.com. Telephone: +34985104787.

# Abstract

# Keywords

*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic grasslands

# Introduction

The mesic meadows of Europe are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045]. At a time they were novel plant communities with a composite flora made up of species from different ecological backgrounds, and during historical times their species composition evolved following changes in human activities and management practices [@RN5035; @RN5045]. Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034]. Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].

Traditional classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997]. For this reason, the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

One of the reasons for the high conservation interest of traditional mesic meadows is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031]. However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe. Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain *ex situ* collections of plant genetic resources [@RN3763]. When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].

Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of the biomass and define the structure of the vegetation. The dispersal unit in Poaceae is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e. modified leaves): the lemma and the palea. Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248]. The dispersal unit is often accompanied by awns or hairs (**Fig. 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290]. Meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273]. Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018]. Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024]. Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229]. Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280]. In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229]. Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].

Next in abundance to Poaceae are the legumes of the family Fabaceae. Fabaceae contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040]. The dispersal unit of most Fabaceae is the seed itself (**Fig. 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). Fabaceae are generally hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073]. In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652]. Buried Fabaceae seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e. physical + physiological) [@RN5037]. However, Fabaceae seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037]. As in Poaceae, seeds of Fabaceae have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273]. Fabaceae seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].

Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039]. They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040]. In many of these families (e.g. Asteraceae, Cyperaceae, Dipsacaceae, Lamiaceae, Polygonaceae, Ranunculaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits. The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig. 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig. 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig. 1h**); hairy in *Knautia* (Dipsacaceae, **Fig. 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). In Apiaceae, it is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig. 1m**), Plantaginaceae (**Fig. 1n**), Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig. 1o**). In Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403]. *Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428]. In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128]. Underdeveloped embryos that need to grow inside the seed before germination (i.e. morphological dormancy) are widespread in Ranunculaceae and Apiaceae [@RN3214; @RN1112]. *Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180]. In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814]. In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112]. *Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226]. Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014], and are able to germinate in the dark [@RN2236].

Although a wealth of studies has accumulated, a synthesis of seed ecology in European mesic meadows is missing. In this article, we review for the first time this topic, combining newly-generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319]. The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries. To test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light we apply Bayesian meta-analysis [@RN5000]. Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyze the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same area.

# Materials and methods

@@

To create a list of representative mesic meadow species for inclusion in our analysis, we used a dataset comprising 118 vegetation relevés (i.e. records of plants species co-occurring in sampling plots) from three Western European regions with well-maintained mesic meadows: 43 relevés from Northern Portugal, 25 from the Cantabrian Mountains of Spain and 50 from the Pyrenees. These relevés contain a sample of mesic meadow diversity, as they were recorded along a major stress gradient related to summer drought [@RN2980]: the Pyrenees and Cantabrian Mountains have a temperate macroclimate, whereas Northern Portugal is transitional between the temperate and Mediterranean macroclimates. Furthermore, meadows from the Pyrenees are closest to the European optimum of mesic meadow vegetation while the Portuguese ones are in suboptimal areas at the limit of the European distribution of temperate meadows [@RN4893]. Finally, the traditional management of meadows is relatively well preserved in these three regions compared to their European context [@RN4998; @RN4999].

Study meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing. Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins. Vegetation sampling took place in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article. As expected, the vegetation of the sampled meadows was dominated by Poaceae and Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots. Other 20 families were recorded, the largest of which was Asteraceae; each of these other families represented less than 10% of the total cover.

Using all the relevés, we calculated the cumulative cover of each species in the entire area. To perform the calculation, first we standardized the cover values of the plots by dividing the cover of each species in each plot by the total plant cover in that plot. Then, for each species, we calculated its total cover in the dataset, by summing its standardized cover values from all the plots. Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values. We removed 208 species with cumulative cover values below 2%, considering them to be rare species that might have been recorded by chance and may not represent the core mesic meadows flora. We used the remaining 116 species as the core list to retrieve seed germination data for this article. The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each). The family with more species in the list was Poaceae (22%), with another 17% belonging to Fabaceae, 15% to Asteraceae, 8% to Apiaceae, and the rest of the families representing less than 5% each.

We also used the relevés as a basis to characterize the preferences of the selected species for three ecological drivers of mesic meadow diversity: cold, summer drought and soil reaction [@RN2980]. For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month). For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5). For each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs). The SNC for any given species and variable is the mean of the environmental variable in all the plots were the species occurs, weighted by species cover in each plot [@RN4709]. The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).

@@

From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016, 2017 and 2018. Seed collection followed the methodology of ENSCONET [-@RN2907]. To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678]. We distributed the seeds on the transparent glass of the scanner, in a 10 x 10 grid. For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds. In the case of the black background, we covered samples with a black box to avoid interference from environmental light. For the white background, we used the scanner cover. We digitized the obtained images and stored them in JPEG format (Joint Photographic Experts Group). We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708]. The program calculates several biometric parameters for each seed on the sample, and among these we choose seed length and width. Additionally, we retrieved species values of seed mass from the Seed Information Database [@RN2987] and of seed number at the individual/ramet level from the LEDA database [@RN2319]. The dataset with the length and width measures is available at GitHub (see Data Availability Statement).

>>>>

---
title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: annals-of-botany.csl
link-citations: yes
always_allow_html: true
---

@@

Eduardo Fernández-Pascual^1,2^ ^+^, Madalena Vaz^3^, Beatriz Morais^2^, Ramón Reiné^4^, Joaquín Ascaso^4^, Elías Afif Khouri^5^, Angelino Carta^6^

^1^ IMIB – Biodiversity Research Institute, University of Oviedo, Mieres, Spain; ^2^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^4^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^5^ Agroforestry Engineering Area, Department of Organisms and Systems Biology, Polytechnic School of Mieres, University of Oviedo, 33600 Mieres, Asturias, Spain; ^6^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy

Running title: Seed ecology of European mesic meadows

^+^ Correspondence: IMIB – Biodiversity Research Institute, University of Oviedo, Edificio de Investigación – 5ª planta, C/ Gonzalo Gutiérrez Quirós s/n
33600 Mieres, Spain. Email: fernandezpeduardo@uniovi.es. Telephone: +34 985103000.

# ABSTRACT (300 words)

* **Background and Aims** European mesic meadows are semi-natural open habitats of high biodiversity and an essential part of European landscapes. Their species-rich seed mixes can be a source of plant material for ecological restoration, urban greening and rewilding. Here, we make a synthesis of the seed ecology of mesic meadows.
* **Methods** We combined our own experimental data with data obtained from databases to create a combined dataset containing `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries. We performed a Bayesian meta-analysis of this dataset to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light. We also used multivariate ordination to check the relationship between seed traits (germination and morphology) and species ecological preferences, and to compare the seed ecology of mesic meadows with that of other herbaceous plant communities from the same area.
* **Key Results** The seed ecology of mesic meadows is characterized by (1) high seed germinability when compared to other herbaceous plant communities; (2) low correspondence between seed traits and species ecological preferences; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae. Poaceae produce many light seeds which respond to gap-detecting germination cues (alternating temperatures and light); Fabaceae produce fewer heavy seeds, which need scarification to break their physical dormancy.
* **Conclusions** High germinability of meadow seeds will reduce their capacity to form persistent seed banks, resulting in dispersal limitations to passive regeneration. For centuries, human activities have shaped the regeneration of meadows, leading to a loss of seed dormancy and decoupling seeds from seasonal cycles, as has been found in many domesticated species. The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration, highlighting the importance of active restoration via seed supply.

# KEYWORDS

*Arrhenatheretalia*, *Arrhenatherion*, Asteraceae, Fabaceae, domestication, hay meadows, mesic grasslands, Poaceae, seed germination, seed morphology, semi-natural, species-rich meadows

# INTRODUCTION

The mesic meadows of Europe are semi-natural open habitats that occupy clearings created by human intervention over a temperate wooded landscape [@RN5033; @RN5045], in sites with moderately fertile and well-drained soils [@RN4698]. Once, these were novel plant communities, with a composite flora made up of species from different ecological backgrounds; during historical times, their species composition further evolved following changes in human events and management practices [@RN5035; @RN5045]. Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034]. Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].

Traditional classification of mesic grasslands emphasises the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997]. For this reason, the European Habitats Directive (92/43/EEC) lists as habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

One of the reasons for the high conservation interest of traditional mesic meadows is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed material to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031]. However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe. Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain *ex situ* collections of plant genetic resources [@RN3763]. When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].

Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of their biomass and define the structure of the vegetation. The dispersal unit in Poaceae is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e. modified leaves): the lemma and the palea. Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248]. The dispersal unit often carries awns or hairs (**Fig. 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290]. Meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273]. Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018]. Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024]. Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be unable to germinate in darkness [@RN1229]. Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280]. In *Poa trivialis*, germination is promoted by light but not by diurnal alternating temperatures (i.e. germination conditions where different temperatures are applied during the day and the night, in diurnal cycles) [@RN1205], although alternating temperatures do encourage some germination in darkness [@RN1229]. Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].

Next in abundance to Poaceae are the legumes of the family Fabaceae. Fabaceae species contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040]. The dispersal unit of most Fabaceae is the seed itself (**Fig. 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). Fabaceae are generally hard-seeded: they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073]. In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652]. Buried Fabaceae seeds can track the seasons, and in some species, germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e. physical + physiological) [@RN5037]. However, Fabaceae seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037]. As in Poaceae, seeds of Fabaceae have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273]. Fabaceae seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].

Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the high biodiversity of the system and increase the aesthetic value of the meadows as perceived by people [@RN5038; @RN5039]. They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040]. In many of these families (e.g. Asteraceae, Cyperaceae, Dipsacaceae, Lamiaceae, Polygonaceae, Ranunculaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits. The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig. 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig. 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig. 1h**); hairy and with an elaiosome in *Knautia* (Dipsacaceae, **Fig. 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). In Apiaceae, it is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families, the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig. 1m**), Plantaginaceae (**Fig. 1n**), Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig. 1o**). In Asteraceae, high germination without previous treatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403]. *Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276] and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428]. In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128]. Underdeveloped embryos that need to grow inside the seed before germination (i.e. morphological dormancy) are widespread in Ranunculaceae and Apiaceae [@RN3214; @RN1112]. *Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages have been reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180]. In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not: this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814]. In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112]. *Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without previous treatment has also been reported [@RN298; @RN226]. Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014] and can germinate in the dark [@RN2236].

Although a wealth of studies has accumulated, a synthesis of the seed ecology of European mesic meadows is still missing. In this article, we review for the first time this topic, combining newly generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319]. The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries. We use this dataset to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light, applying Bayesian meta-analysis [@RN5000]. Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyse the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same geographic area.

# MATERIALS AND METHODS

@@

To create a list of representative mesic meadow species for inclusion in our analysis, we used a dataset comprising 118 vegetation relevés (i.e. records of plants species co-occurring in sampling plots) from three Western European regions with well-maintained mesic meadows: 43 relevés from Northern Portugal, 25 from the Cantabrian Mountains of Spain and 50 from the Pyrenees. These relevés contain a sample of mesic meadow diversity, as they were recorded along a major stress gradient related to summer drought [@RN2980]: the Pyrenees and Cantabrian Mountains have a temperate macroclimate, whereas Northern Portugal is transitional between the temperate and Mediterranean macroclimates. Furthermore, meadows from the Pyrenees are closest to the European optimum of mesic meadow vegetation while the Portuguese ones are in suboptimal areas at the limit of the European distribution of temperate meadows [@RN4893]. Finally, the traditional management of meadows (i.e. mowing for haymaking once or twice per year plus light grazing) is relatively well preserved in these three regions compared to their European context [@RN4998; @RN4999].

The studied meadows were maintained by traditional agricultural practices: mowing for haymaking once or twice per year plus light grazing. Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of the meadow, avoiding the margins. Vegetation sampling took place in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article. As expected, the vegetation of the sampled meadows was dominated by Poaceae and Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots. Other 20 families were recorded, the largest of which was Asteraceae; each of these other families represented less than 10% of the total cover.

Using all the relevés, we calculated the cumulative cover of each species in the entire area. To perform the calculation, first we standardized the cover values of the plots by dividing the cover of each species in each plot by the total plant cover in that plot. Then, for each species, we calculated its total cover in the dataset, by summing its standardized cover values from all the plots. Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values. From the resulting list of species, we removed 208 species with cumulative cover values below 2%, considering them to be transient species [@RN4977] that might have been recorded by chance and may not represent the core mesic meadows flora. We used the remaining 116 species as the core list of meadows species to retrieve seed data for this article. Most of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each). The family with more species in the list was Poaceae (22%), with another 17% belonging to Fabaceae, 15% to Asteraceae, 8% to Apiaceae, and the rest of the families representing less than 5% each.

## Species ecological requirements

We also used the relevés as a basis to characterize the preferences of the selected species for three environmental factors (cold, summer drought and soil reaction) which have been found to be major ecological drivers of mesic meadow plant diversity [@RN2980]. For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month). For soil reaction (pH), we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil:water (1:2.5). With each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs). The SNC for any given species and variable is the mean of the environmental variable in all the plots where the species occurs, weighted by species cover in each plot [@RN4709]. The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).

@@

From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016, 2017 and 2018. Seed collection followed the methodology of ENSCONET [-@RN2907]. To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678]. We distributed the seeds on the transparent glass of the scanner, in a 10 x 10 grid. For each sample, and without moving the seeds, we repeated the scan with black and white backgrounds. In the case of the black background, we covered samples with a black box to avoid interference from environmental light. For the white background, we used the scanner cover. We digitized the obtained images and stored them in JPEG format (Joint Photographic Experts Group). We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708]. The program calculates several biometric parameters for each seed on the sample, and among those we chose seed length and width. Additionally, we retrieved species values of seed mass from the Seed Information Database [@RN2987] and of seed number at the individual/ramet level from the LEDA database [@RN2319]. The dataset with the length and width measures is available at GitHub (see Data Availability Statement).

We also germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, for each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel. For the rest of the families, we used gibberellic acid GA~3~ (0.0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12-hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were mouldy. We only considered normal seeds when calculating germination proportions and conducting subsequent analyses.
